---
title: "R4DS 27_6 YAML header"
author: "Gabriel"
date: "12/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# YAML Headers

You can control the whole document by adjusting the YAML headers

## Parameters

Parameters are useful because they let you re-render entire documents with a key variable changed. This could be useful if you wanted to generate a report for say, each student in a class.

The params field lets you set parameters

For example, the following setup would let you create a report for each class of car quite easily
---
output: html_document
params:
  my_class: "suv"
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

class <- mpg %>% filter(class == params$my_class)
```

### Fuel economy for `r params$my_class`s

```{r, message = FALSE}
ggplot(class, aes(displ, hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

You can also run R expressions as parameters if you put !r first, this can be helpful for things like setting dates and times:

params:
  start: !r lubridate::ymd("2015-01-01")
  snapshot: !r lubridate::ymd_hms("2015-01-01 12:30:00")

RStudio has the option to "Knit with Parameters" in the Knit dropdown menu. This is a user-friendly way to generate reports with different parameters.
In code, you can call a function to do it: 

```{r}
rmarkdown::render("fuel-economy.Rmd", params = list(my_class = "suv"))
```

This can be combined with purrr::pwalk(), for example to generate a report for each value of class in mpg:

```{r}
# Create a data frame with one row for each class, with the filename and the params
reports <- tibble(
  class = unique(mpg$class),
  filename = stringr::str_c("fuel-economy-", class, ".html"),
  params = purrr::map(class, ~ list(my_class = .))
)
reports

# Then use purrr's parallel walk to call render once for each row
reports %>%
  select(output_file = filename, params) %>% 
  purrr::pwalk(rmarkdown::render, input = "fuel-economy.Rmd")
```

## Bibliographies and Citations
Pandoc can generate citations and bibliographies in several styles. First you specify a bibliography file, containing the paths from the directory that contains the .Rmd file:

bibliography: rmarkdown.bib

Then to create a citation you use @ + the citation identifier you've setup in the bibliography file.

Inside brackets might look like this:
Separate multiple citations with a `;`: Blah blah [@smith04; @doe99].

You can add arbitrary comments inside the square brackets: 
Blah blah [see @doe99, pp. 33-35; also @smith04, ch. 1].

Remove the square brackets to create an in-text citation: @smith04 
says blah, or @smith04 [p. 33] says blah.

Add a `-` before the citation to suppress the author's name: 
Smith says blah [-@smith04].

When the file is rendered, it will build and append a bibliography. The one thing it doesn't do is create a heading, so it's good to end it with # References

You can change the citation stle by referencing a citation style language (CSL). This is a file which you also have to specify the path to.

bibliography: rmarkdown.bib
csl: american-physiological-society.csl 

Common CSLs can be found here: http://github.com/citation-style-language/styles.

**There's a way to integrate this with Zotero but I don't know how yet**
https://christopherjunk.netlify.app/blog/2019/02/25/zotero-rmarkdown/
https://gsverhoeven.github.io/post/zotero-rmarkdown-csl/
https://dewey.dunnington.ca/post/2020/getting-started-zotero-better-bibtex-rmarkdown/

# Learning more

R Markdown is young and growing rapidly, the best place to keep up is http://rmarkdown.rstudio.com

He hasn't touched at all on collaberating and style

Collaberation often involves version control, and he recommends two places to learn git:
- Happy Git with R http://happygitwithr.com
- Git and Github chapter of R Packages http://r-pkgs.had.co.nz/git.html

For style he's recommend either (or both!):
- _Style: Lessons in Clarity and Grace_ by Joseph M. Williams & Joseph Bizup
- _The Sense of Structure: Writing from the Reader’s Perspective_ by George Gopen

Both can probably be found relatively cheaply second-hand
George Gopen also has some articles, aimed at lawyers but also applies to data scientists https://www.georgegopen.com/the-litigation-articles.html