---
title: "R4DS 27_5 Troubleshooting"
author: "Gabriel"
date: "12/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Troubleshooting can be tricky because you're not in an interactive environment anymore

First you should try to recreate the problem, restarting R and running all chunks again (Ctrl + Alt + R). Hopefully that will recreate the issue and then you'll be able to use the interactive R console to figure out what was going wrong.

Otherwise, issues might be between the interactive environment and the markdown environment.

The working directory for markdown files is the directory it's in. You can check it's what you expect with getwd()

```{r}
getwd()
```

Then you'd systematically go through things that might be causing the bug, and that're different between your interactive and markdown session

The easiest way to do that is to set error = TRUE on the chunk and then use print() and str() to check if everything makes sense.

