---
title: "Markdown Formats"
author: "Gabriel"
date: "2023-01-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Markdown Formats

We've done HTML, but markdown can create other types of output
You can adjust this permenantly by modfying the YAML header

Or by calling the render by hand:

```{r}
rmarkdown::render("diamond-sizes.Rmd", output_format = "word_document")
```

Or using the drop-down menu in RStudio

## Output options

Each output format is associated with a function. This can help you know how to set parameters. For example, the html_document parameters can be found with `?rmarkdown::html_document`.

To override the default parameters, you need to expand the output field. This is what you'd do to make a html with a floating table of contents:

```{r}
output:
  html_document:
    toc: true
    toc_float: true
```

Or you could also output to multiple formats

```{r}
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document: default
```

But here you have to specify that you want the default options on the pdf.

## Documents

There are a range of default document types in R.

- `html_document` makes an HTML as we've seen
- `pdf_document` makes a PDF
- `word_document` makes a MS Word document (.docx)
- `rtf_document` for Rich Text Format (.rtf)
- `md_document` for a Markdown file
- `github_document` for a Markdown file specifically tailored for sharing on Github

When sharing with others who don't need to see the code, you can turn off the default display of code in the setup chunk

```{r}
knitr::opts_chunk$set(echo = FALSE)
```

HTML documents can also have hidden code that is revealed with a click

```{r}
output:
  html_document:
    code_folding: hide
```


## Notebooks

`html_notebook` is a variation on the HTML document. It's made for collaberating with others, so it contains the full source code. Unlike `html_document` which is made for sharing with decision makers, and so doesn't contain the code.

Notebooks can be viewed in the web browser to view the output. Or it can be opened in RStudio, which recreates the .Rmd file that generated it.

If using git and Github he recommends these as default outputs:

```{r}
output:
  html_notebook: default
  github_document: default
```


## Presentations

Markdown can create simple presentations that can save a lot of time.

Content is divided into slides, with each slide beginning with each first `#` or second `##` level header. Or you can add a new slide with no header using `***`

There are 3 built in presentation formats:
1. `ioslides_presentation` - HTML presentation with ioslides
2. `slidy_presentation` - HTML presentation with W3C Slidy
3. `beamer_presentation` - PDF presentation with LaTeX Beamer

But there are packages with more options


## Dashboards

These can convey a lot of information quickly. Generally:
- Each level 1 header begins a new page
- Each level 2 header begins a new column
- Each level 3 header begins a new row

Flexdashboard has tools for side-bars, gagues, value boxes and other things. http://rmarkdown.rstudio.com/flexdashboard/


## Interactivity

HTML documents can contain interactive components

### htmlwidgets

An example here is the leaflet map, which lets you build an interactive map

```{r}
library(leaflet)
leaflet() %>%
  setView(174.764, -36.877, zoom = 16) %>% 
  addTiles() %>%
  addMarkers(174.764, -36.877, popup = "Maungawhau") 
```

And the package handles all the details. Others include:
- *dygraphs*, http://rstudio.github.io/dygraphs/, for interactive time series visualisations
- *DT*, http://rstudio.github.io/DT/, for interactive tables
- *threejs*, https://github.com/bwlewis/rthreejs for interactive 3d plots
- *DiagrammeR*, http://rich-iannone.github.io/DiagrammeR/ for diagrams (like flow charts and simple node-link diagrams)

More can be found at  http://www.htmlwidgets.org/


### Shiny
htmlwidgets are interactive on the client side. It is built on what can be done in HTML and JavaScript

shiny lets you use R to create interactivity. To do this you need to add `shiny` to the header

```{r}
title: "Shiny Web App"
output: html_document
runtime: shiny
```

Then you can use the input functions to add interactive components

```{r}
library(shiny)

textInput("name", "What is your name?")
numericInput("age", "How old are you?", NA, min = 0, max = 150)
```

Then you can refer to the values with `input$name` and `input$age` so the code can update each time it's run.

Unfortunately, Shiny apps need a Shiny server to be run online. Running the apps on your own machine creates a server for you on the machine. But to share it you need a public server.

Learn more about Shiny at http://shiny.rstudio.com/


## Websites

This can be done with only a little extra infrastructure:
- Put the `.Rmd` fules in a single directory. `index.Rmd`. This will become the home page
- Add a YAML file named `_site.yml` which provides the navigation to the site. E.g.:

```{r}
name: "my-website"
navbar:
  title: "My Website"
  left:
    - text: "Home"
      href: index.html
    - text: "Viridis Colors"
      href: 1-example.html
    - text: "Terrain Colors"
      href: 3-inline.html
```

Execute `rmarkdown::render_site()` to build the `_site` which is a set of files which can be turned into a static website. Or if it's in an RStudio Project then RStudio will add a Build tab so you can build and preview your site.

More at http://rmarkdown.rstudio.com/rmarkdown_websites.html


## Other formats

- The bookdown package, https://github.com/rstudio/bookdown, makes it easy to write books, like this one. To learn more, read Authoring Books with R Markdown, by Yihui Xie, which is, of course, written in bookdown. Visit http://www.bookdown.org to see other bookdown books written by the wider R community
- The prettydoc package, https://github.com/yixuan/prettydoc/, provides lightweight document formats with a range of attractive themes
- The rticles package, https://github.com/rstudio/rticles, compiles a selection of formats tailored for specific scientific journals

## Learning more

- To improve your presentation skills, I recommend Presentation Patterns, by Neal Ford, Matthew McCollough, and Nathaniel Schutta. It provides a set of effective patterns (both low- and high-level) that you can apply to improve your presentations
- If you give academic talks, I recommend reading the Leek group guide to giving talks
- If you are creating a lot of dashboards, make sure to read Stephen Few’s Information Dashboard Design: The Effective Visual Communication of Data. It will help you create dashboards that are truly useful, not just pretty to look at
- Effectively communicating your ideas often benefits from some knowledge of graphic design. The Non-Designer’s Design Book is a great place to start